Assessing the Population-level Conservation Effects of Marine Protected Areas

### 2021-03-14 documents/ovando-et-al-pop-effects-of-mpas.Rmd
Minor revisions (#6)
* addressing Owen's minor edits and figure improvements
* merging back to main

Co-authored-by: Owen R. Liu <owenrliu@gmail.com>

---
title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
abstract: | 
  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders. We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically. A network of MPAs covering roughly 20% of the Channel Islands National Marine Sanctuary was put in place in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach. We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect. Clearly communicating expectations and uncertainties around MPAs is critical to ensuring that MPAs play effective roles marine conservation. Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
bibliography: references.yaml
csl: conservation-biology.csl
output:
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: xelatex
  bookdown::html_document2: default
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
toc: false
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{draftwatermark}
  - \SetWatermarkText{DRAFT}
always_allow_html: true
---

@@

theme_2col <- hrbrthemes::theme_ipsum(base_size = 10,
                                        axis_title_size = 12,
                                      plot_title_size = 12)

@@

where $d_i$ is the biomass density at observation *i*. To account for the fact that MPA effects will evolve over time, we estimate a vector of MPA effects in three-year blocks for the all years after the MPAs were implemented in 2003. **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water temperature and observer experience. **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.  Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment ($MPA=1$) on the treated targeted group ($T=1$) (Table\@ref(tab:coef-table)). We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically. Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].

@@

```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs. Panel (a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (y-axis), and pre-MPA depletion (x-axis). Pre-MPA depletion is a measure of fishing pressure, where 0% means that the population is unfished, and 100% means that the population is extinct in the time period immediately prior to MPA implementation. Panel (b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}

>>>>

---
title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
abstract: | 
  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader populations that span *beyond* their borders. We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically. A network of MPAs covering roughly 20% of the Channel Islands National Marine Sanctuary was put in place in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach. We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect. Clearly communicating expectations and uncertainties around MPAs is critical to ensuring that MPAs play effective roles marine conservation. Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
bibliography: references.yaml
csl: conservation-biology.csl
output:
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: xelatex
  bookdown::html_document2: default
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
toc: false
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{draftwatermark}
  - \SetWatermarkText{DRAFT}
always_allow_html: true
---

@@

theme_2col <- hrbrthemes::theme_ipsum(base_size = 10,
                                        axis_title_size = 14,
                                      plot_title_size = 12)

@@

where $d_i$ is the biomass density at observation *i*. To account for the fact that MPA effects will evolve over time, we estimate a vector of MPA effects in three-year blocks for the all years after the MPAs were implemented in 2003. **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water temperature and observer experience. **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.  Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment ($MPA=1$) on the treated targeted group ($T=1$) (Table\@ref(tab:coef-table)). We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically. Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals (also termed "uncertainty" or "compatibility" intervals) [@gelman2014a;@mcelreath2020];.

@@

```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs. Panel (a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (y-axis), and pre-MPA depletion (x-axis). Pre-MPA depletion is a measure of fishing pressure, where 0% means that the population is unfished, and 100% means that the population is extinct in the time period immediately prior to MPA implementation. Panel (b) shows distribution of simulations across range of MPA sizes and pre-MPA depletions separately."}

### 2021-03-14 documents/pop-effects-of-mpas-front-matter.Rmd
Minor revisions (#6)
* addressing Owen's minor edits and figure improvements
* merging back to main

Co-authored-by: Owen R. Liu <owenrliu@gmail.com>

The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to detect

@@

@@

5976

@@

Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship. Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant. This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to K. Davis Koehn and A. Parsons-Field. We also thank C. Szuwalski, J. Lawson, A. Punt, Chris Brown, and two anonymous reviewers for helpful comments and technical support.

>>>>

The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to measure and detect

@@

Channel Islands National Marine Sanctuary

@@

5994

@@

Funding for the simulation model was provided by the National Marine Fisheries Service-SeaGrant Population and Ecosystem Dynamics Fellowship. Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant. This study would not have been possible without the work provided by PISCO divers over the years, with special thanks to K. Davis Koehn and A. Parsons-Field. We also thank C. Szuwalski, J. Lawson, A. Punt, C. Brown, and two anonymous reviewers for helpful comments and technical support.
